$\pi$ with leftovers: \\ a mechanisation in Agda

### 2021-01-11 paper/paper.tex
Shorten the intro to the type system

We use \emph{leftover typing} \cite{Allais2018a} for our type system, an approach that, in addition to the usual typing context $\PreCtx_n$ and (input) usage context $\Ctx_{idxs}$, adds an extra \emph{(output) usage context} $\Ctx_{idxs}$ to the typing rules.
This output context contains the \emph{leftovers} (the unused multiplicities) of the process being typed.
These leftovers can then be used as an input to another typing derivation.

Typing judgments for linear calculi typically carry a single context of usage annotations and thus the typing rules require frequent top-down proofs of context splitting --- e.g., the typing rule for parallel composition would look like:
\todo{we have this in th intro already}
\begin{mathpar}
  \inferrule
  {\opctx{\Gamma}{\Delta}{\Xi} \\ \Delta \; \type{\vdash} \; P \\ \Xi \; \type{\vdash} \; Q}
  {\Gamma \; \type{\vdash} \; \comp{P}{Q}}
\end{mathpar}
Input requires a similar context split; output requires two such context splits.
This is rather cumbersome: for input and output, the proof of context split carries \emph{exactly} the same information as the typing rule for variables; for parallel composition it means that the typing derivations of both subprocesses will have no context information whatsoever until the context split is determined.

Leftover typing solves this problem by inverting the flow of information so that \textbf{the typing derivations of the subprocesses determine the context split}.
As a result, usage information is threaded bottom-up through the typing derivation, and top-down context split proofs are no longer needed.
In addition, it allows \emph{framing} (\autoref{thm:framing}) to be stated, and \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) to acquire a more general form.

Our type system with leftovers is composed of two typing relations: one for variable references (\autoref{def:varref}) and one for processes (\autoref{def:types}).
Both relations are indexed by a typing context $\gamma$, an input usage context $\Gamma$, and an output usage context $\Delta$ (the leftovers).

>>>>

We present a resource-aware type system for the \picalc{} based on \emph{leftover typing} \cite{Allais2018a}, a technique that, in addition to the usual typing context $\PreCtx_n$ and (input) usage context $\Ctx_{idxs}$, adds an extra \emph{(output) usage context} $\Ctx_{idxs}$ to the typing rules.
This output context contains the \emph{leftovers} (the unused multiplicities) of the process being typed.
These leftovers can then be used as input to another typing derivation.

Leftover typing inverts the information flow concerning usage annotations so that it is the typing derivations of subprocesses which determine how resources are allocated.
As a result, context split proofs are no longer necessary.
Leftover typing also allows \emph{framing} (\autoref{thm:framing}) to be stated, and \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) to acquire a more general form.

Our type system is composed of two typing judgments: one for variable references (\autoref{def:varref}) and one for processes (\autoref{def:types}).
Both judgments are indexed by a typing context $\gamma$, an input usage context $\Gamma$, and an output usage context $\Delta$ (the leftovers).
